---
title: "Data Manipulation"
output: github_document
---

```{r setup}
library(tidyverse)

# tidyverse is collection of packages
```

## Load in the FAS Litters Data

```{r}
litters_df = read_csv("./data/data/FAS_litters.csv")

# we are going to update variable names to make them a bit easier

litters_df = janitor::clean_names(litters_df)

litters_df
```

## `select`

choose some columns and not others

```{r}
select(litters_df, group, litter_number)

select(litters_df, group, gd0_weight)

select(litters_df, group, gd0_weight:gd_of_birth)

# when using select(), you first include the data you wanna use and then the columns you want to keep
# use colon as "range"

```

```{r}
select(litters_df, -litter_number)

# this will get rid of only a specific column and keep everything else in data set

```

renaming columns

```{r}
select(litters_df, GROUP = group, LITTer_NUmber = litter_number)
```
```{r}
rename(litters_df, GROUP = group, LITTer = litter_number)

# can use select but rename is also an available function to use
```

select helper

```{r}
# using some additional functions - selecting columns that we know we want to keep based on some criteria.

select(litters_df, starts_with("gd"))
```

```{r}
select(litters_df, litter_number, everything())

# will move litter_number from wherever it is to the beginning and keep everything else
```

```{r}
relocate(litters_df, litter_number)

# does the same thing as above; helpful to know it exists
```

## filter

operates on rows - intended to remove/keep rows that you're interested in for analysis 

need to give it a condition 


the real challenge in using filter is finding how you might specify or coming up with a way to specify the collection of rows that you actually want to keep

```{r}
filter(litters_df, gd0_weight < 22)

# filtering rows based on whether or not this observation is true
```

```{r}
filter(litters_df, gd0_weight >= 22)
```

```{r}
filter(litters_df, gd_of_birth == 20)

# the two equal signs is a test for equality
```

```{r}
filter(litters_df, !(gd_of_birth == 20))
       
# create logical operation that says see where this is true and not true, and then flip both of those

filter(litters_df, (gd_of_birth != 20))

# "!=" -> represents "not equal to"
```

```{r}
filter(litters_df, gd0_weight >= 22, gd_of_birth == 20)
```

```{r}
filter(litters_df, group == "Mod8")
```

```{r}
filter(litters_df, group %in% c("Con7", "Mod8"))

# shows where group is either of those two options -> either Con7 or Mod8, here are the rows where that shows to be true
```

## mutate

mutate will allow us to create additional values -> neither modify an existing variable or create some new variable

```{r}
mutate(
  litters_df, wt_gain = gd18_weight - gd0_weight, 
  group = str_to_lower(group))

# modify the group variable by making everything lowercase and simulataneously override the group variable with a modified version
```

## `arrange`

arrange - goal is to put things in order

```{r}
arrange(litters_df, pups_born_alive)

# arranging data frame by pups_born_alive
```

##`%>%` pipe operation

```{r}
# what would happen if we wanted to import a dataset, then clean the names, select some variables, then mutate step, then filter

litters_data_raw = read_csv("./data/data/FAS_litters.csv")

litters_clean_name = janitor::clean_names(litters_data_raw)
litters_data_selected = select(litters_clean_name, -pups_survive)
litters_mutated = mutate(litters_data_selected, wt_gain = gd18_weight - gd0_weight)
litters_without_missing = drop_na(litters_mutated, gd0_weight)

# we are going to drop anything where data is missing (above shows where gd0_weight is missing)
```

USE THE PIPE OPERATOR INSTEAD

```{r}
litters_df = 
  read_csv("./data/data/FAS_litters.csv") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(wt_gain = gd18_weight - gd0_weight) %>%
  drop_na(gd0_weight)

# we don't need to put anything in the parantheses because the "pipe" says whatever your first argument was is going to continue (don't need to repeat)

# this code chunk shows the "pipes" versio of the code chunka above to sling together a chain of commands
```

